---
title: "Student performance in exam"
author: "Donggyu Isaac Hills"
date: '2021 3 1 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(janitor)
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(caTools)


exam <- read.csv(file = "StudentsPerformance.csv")
```

## Set up 
```{r set up}
# Checking if there is any missing values, which returns none
which(is.na(exam), arr.ind=TRUE)
# Renaming columns
colnames(exam)<-gsub("\\.","\\_",colnames(exam))
# Shows changes
head(exam, 3)
# Summary of data
summary(exam)
```

## Does gender affect performance?
```{r gender}
num_gender <- tabyl(exam, gender)

plot_gender <- ggplot(data = exam, 
                      aes(x = gender,
                          fill = gender)) + 
               geom_bar(stat = "count") +
               xlab("Gender") + 
               ylab("Count") +
               ggtitle("Male Female Ratio")

mean_gender <- aggregate(exam[, 6:8], by = list(exam$gender), FUN = mean, na.rm = T)

# Comparing math score by gender  
gender_math_graph <- exam %>% select(gender, `math_score`) %>% filter(!is.na(math_score)) %>% ggplot(aes(x=gender, y = math_score, color = gender)) + geom_boxplot() + geom_jitter(width = 0.2, alpha = 0.2)


# Comparing reading score by gender 
gender_reading_graph <- exam %>% select(gender, `reading_score`) %>% filter(!is.na(reading_score)) %>% ggplot(aes(x=gender, y = reading_score, color = gender)) + geom_boxplot() + geom_jitter(width = 0.2, alpha = 0.2)


#Comparing writing score by gender 
gender_writing_graph <- exam %>% select(gender, `writing_score`) %>% filter(!is.na(writing_score)) %>% ggplot(aes(x=gender, y = writing_score, color = gender)) + geom_boxplot() + geom_jitter(width = 0.2, alpha = 0.2)
```

```{r gender graph , echo=FALSE}
## Does gender affect performance?
num_gender
plot_gender
mean_gender 
gender_math_graph
gender_reading_graph
gender_writing_graph
```

Analysis  <br>1. 518 female and 482 male students.<br>
          <br>2. Female students performed better on reading and writing     
           section.<br>
          <br>3. Based on this dataset, a correlation between students' performance 
           and their gender is quite unclear.<br>

## Does race/ethinicity affect performance?
```{r Race}
num_race <- tabyl(exam, race_ethnicity)

plot_race <- ggplot(data = exam, 
                      aes(x = race_ethnicity,
                          fill = race_ethnicity)) + 
               geom_bar(stat = "count") +
               xlab("Race") + 
               ylab("Count") +
               ggtitle("Race Ratio")

mean_race <- aggregate(exam[, 6:8], by = list(exam$race_ethnicity), FUN = mean, na.rm = T)

math_race = ggplot(data = exam) + 
  geom_boxplot(mapping = aes(x = race_ethnicity, y = math_score, fill = race_ethnicity)) +
  labs(x = "Race/Ethnicity", y = "Math Scores")


reading_race = ggplot(data = exam) + 
  geom_boxplot(mapping = aes(x = race_ethnicity, y = reading_score, fill = race_ethnicity)) +
  labs(x = "Race/Ethnicity", y = "Reading Scores")
  

writing_race = ggplot(data = exam) +
  geom_boxplot(mapping = aes(x = race_ethnicity, y = writing_score, fill = race_ethnicity)) +
  labs(x = "Race/Ethnicity", y = "Writing Scores")
```

```{r Race graph, echo=FALSE}
num_race
plot_race
mean_race 
math_race
reading_race
writing_race
```


Analysis   <br>1. This dataset contains 89 of group A, 190 of group B,319 of group C
           , 262 of group D, and 140 of group E.<br>
           <br>2. In overall, While group E performed the best, group A performed 
           the worst.<br>
           <br>3. Based on this dataset, there is a positive correlation between 
           students' performance and their racial background/ethnicity.<br>
           
## Does parental education affect performance? 
``` {r Parents education}
num_parent <- tabyl(exam, parental_level_of_education)

plot_parental_level_of_education <- ggplot(data = exam, 
                      aes(x = parental_level_of_education,
                          fill = parental_level_of_education)) + 
               geom_bar(stat = "count") +
               xlab("Parental level of education") + 
               ylab("Count") +
               ggtitle("Parental level of education Ratio")

mean_parent <- aggregate(exam[, 6:8], by = list(exam$parental_level_of_education), FUN = mean, na.rm = T)

math_parent = ggplot(data = exam) + 
  geom_boxplot(mapping = aes(x = parental_level_of_education, y = math_score, fill = parental_level_of_education)) +
  labs(x = "Parent Education", y = "Math Scores")


reading_parent = ggplot(data = exam) + 
  geom_boxplot(mapping = aes(x = parental_level_of_education, y = reading_score, fill = parental_level_of_education)) +
  labs(x = "Parent Education", y = "Reading Scores")
  

writing_parent = ggplot(data = exam) +
  geom_boxplot(mapping = aes(x = parental_level_of_education, y = writing_score, fill = parental_level_of_education)) +
  labs(x = "Parent Education", y = "Writing Scores")
 

```

```{r parent graph, echo=FALSE}
num_parent
plot_parental_level_of_education
mean_parent
math_parent
reading_parent
writing_parent
```


Analysis   <br>1. In this dataset, there are 222 of associate's degree, 118 of
           bachelor's degree, 196 high school, 59 of master's degree, 226 of
           some college, 179 of some high school as parental level of education.<br> 
           <br>2. In overall, while the students whose parents have master degree 
           performed the best, the students whose parents have high school 
           performed the worst.<br>
           <br>3. Based on this dataset, there is a positive correlation between 
           students' performance and their parental level of education.<br>

## Does type of lunch affect performance?
```{r lunch type}
num_lunch <- tabyl(exam, lunch)

plot_lunch <- ggplot(data = exam,
                aes(x = lunch, fill = lunch)) +
         geom_bar(stat = "Count")

lunch_stat <- aggregate(exam[, 6:8], by = list(exam$lunch), FUN = mean, na.rm = T)

lunch_graph_math = ggplot(data = exam) + 
  geom_boxplot(mapping = aes(x = lunch, y = math_score, fill = lunch)) +
  labs(x = "type of lunch", y = "Math Scores")

lunch_graph_reading = ggplot(data = exam) + 
  geom_boxplot(mapping = aes(x = lunch, y = reading_score, fill = lunch)) +
  labs(x = "type of lunch", y = "Reading Scores")

lunch_graph_writing = ggplot(data = exam) + 
  geom_boxplot(mapping = aes(x = lunch, y = writing_score, fill = lunch)) +
  labs(x = "type of lunch", y = "writing Scores")



```

```{r lunch graph, echo=FALSE}
num_lunch
plot_lunch
lunch_stat
lunch_graph_math
lunch_graph_reading
lunch_graph_writing
```


Analysis   <br>1. In this dataset, there are 355 students who receive free/reduced 
           lunch and 645 students who pay regular amount for lunch.<br>
           <br>2. Though the students who have standard lunch type are almost double
           of the students who receive free/reduced lunch, they performed much 
           better than the students' with reduced/free lunch in overall.<br>
           <br>3. Based on this dataset, there is a correlation between students' 
           performance and their lunch type.<br>
           
```{r Linear}
#Predicting the reading score by using writing score
cor(exam$reading_score,exam$writing_score)

score <- exam[,-2:-5]

ggplot(exam,aes(x=reading_score,y=writing_score))+
geom_point()+
stat_smooth(method= "lm",col="blue")

#Split data set into 2 
split <- sample.split(score,SplitRatio = 0.75)
train <- subset(score, split="TRUE")
test <- subset(score, split="FALSE")

#Linear Model
lm <- lm(exam$reading_score~exam$writing_score,data=train)
summary(lm)

#Predict the reading score
reading_predict <- predict(lm,train)

#Predict reading data vs Actual reading data 
plot(exam$reading_score,type="l",lty=1.8,col="red")
lines(reading_predict,type="l", col="blue")

#predict the reading score using math score
cor(exam$reading_score,exam$math_score)

ggplot(exam,aes(x=reading_score,y=math_score))+
geom_point()+
stat_smooth(method= "lm",col="blue")

lm_df <- lm(score$reading_score~score$math_score,data=train)
summary(lm_df)

reading_predict_2 <- predict(lm_df,train)

#predicted reading score vs actual reading score
plot(score$reading_score,type="l",col="red")
lines(reading_predict_2,type="l",col="blue")
```
<br> Successfully predicted reading score by using writing and math score.<br>
<br> The predicted score is very close to the actual data. <br> 
